Data Provenance for Concurrent Program Debugging and Analysis in Production Systems

### 2013-11-10 paper.tex
committing some changes before sending a draft out

Data provenance information is directly useful during program debugging.  prior
work~\cite{whylinechi,whylineicse} community showed that when programmers debug
programs, they benefit from the ability to directly ask such provenance
(``why'') questions of their debugger.  Origin tracking~\cite{badapples} and
program slicing~\cite{tipslicingsurvey} also make debugging easier by reporting
the provenance of values.  Like these prior efforts, in this work we focus on
using data provenance information for {\em debugging} -- understanding the root
cause of a failure -- as opposed to {\em bug detection} -- finding new, unknown
bugs.

@@

When the failure manifests, Thread 1 is at the assertion in {\tt append()}, as
picture.  However, Thread 2 may have completed its erase operation and moved on
to execute some unrelated code.  The call stack for Thread 2 may show nothing
about its involvement in the failure, despite the fact that its interleaving
with the {\tt append()} code is the root cause.  Examining the core dump, the
programmer can see that the length and buffer are inconsistent, but is left to
ask {\em why} they are inconsistent.

@@

{\bf TODO: Replace with the "fixed" version of the JDK figure (currently in eval).  Update the text to match the new figure.}

@@

The running example in Figures~\ref{fig:coreDumpFail} and ~\ref{fig:httlws},
and the text in Section~\ref{sec:debugging} describes the benefit provided by
last writer slices in debugging an ordering violation in HTTrack-3.43.9.

{\bf TODO: Move this to Section 3 where it will live.}
\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{figs/JDKStringBufferDebug.pdf}
\caption{\label{fig:jdklws}{\bf Using last writer slices to debug an
ordering violation in the JDK 1.4 StringBuffer library.} The last writer slice reveals the interaction between the erase function and the append function that leads to a failure.}
\end{figure}

>>>>

Data provenance information is directly useful during program debugging.  prior
work~\cite{whylinechi,whylineicse} showed that when programmers debug
programs, they benefit from the ability to directly ask such provenance
(``why'') questions of their debugger.  Origin tracking~\cite{badapples} and
program slicing~\cite{tipslicingsurvey} also make debugging easier by reporting
the provenance of values.  Like these prior efforts, in this work we focus on
using data provenance information for {\em debugging} -- understanding the root
cause of a failure -- as opposed to {\em bug detection} -- finding new, unknown
bugs.

@@

When the failure manifests, Thread 1 is at the assertion in {\tt append()}, as
pictured.  However, Thread 2 may have completed its erase operation and moved on
to execute some unrelated code.  The call stack for Thread 2 may show nothing
about its involvement in the failure, despite the fact that its interleaving
with the {\tt append()} code is the root cause.  Examining the core dump, the
programmer can see that the length and buffer are inconsistent, but is left to
ask {\em why} they are inconsistent.

@@

{\bf TODO: Update the text to explain how the last writer slice shown makes the conection between erase() and append(), showing the programmer their mistake.}

\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{figs/JDKStringBufferDebug2.pdf}
\caption{\label{fig:jdklws}{\bf Using last writer slices to debug an
ordering violation in the JDK 1.4 StringBuffer library.} The last writer slice reveals the interaction between the erase function and the append function that leads to a failure.}
\end{figure}

@@

The running example in Figures~\ref{fig:coreDumpFail} and ~\ref{fig:httlws},
and the text in Section~\ref{sec:debugging} describes the benefit provided by
last writer slices in debugging an ordering violation in the JDK-1.4 stringbuffer library.
